Sequential and Parallel Solution-Biased Search for Subgraph Algorithms

### 2019-06-01 talk/talk.tex
A talk

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}
\usepackage{graphicx}

@@

        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

@@

@@

\title[Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms]{Value Ordering,
Discrepancies, and Restarts for Subgraph Algorithms}
\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick
Prosser and \\ James Trimble}

@@

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, width=\paperwidth]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\section{Subgraph Isomorphism}

\begin{frame}{Non-Induced Subgraph Isomorphism}

    \begin{center}\begin{tikzpicture}
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);

        \draw <2> [ultra thick] (Na) -- (Nb);
        \draw <2> [ultra thick] (Nb) -- (Nc);
        \draw <2> [ultra thick] (Nc) -- (Nd);
        \draw <2> [ultra thick] (Nb) -- (Nd);

        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <2> [ultra thick] (N1) -- (N3);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <2> [ultra thick] (N3) -- (N5);
        \draw <2> [ultra thick] (N3) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <2> [ultra thick] (N5) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}\end{center}

\end{frame}

\begin{frame}{Benchmark Instances}

    \begin{itemize}
        \item Other papers: a subset of 5,725 instances from Christine Solnon's selection:
            \begin{itemize}
                \item Randomly generated with different models.
                \item Real-world graphs.
                \item Computer vision problems.
                \item Biochemistry problems.
                \item Phase transition instances.
            \end{itemize}
        \item Here: all 14,621 instances:
            \begin{itemize}
                \item $\ge \input{../paper/gen-n-sat.tex}$ satisfiable.
                \item $\ge \input{../paper/gen-n-unsat.tex}$ unsatisfiable.
            \end{itemize}
        \item A lot of them are very easy for good algorithms.
    \end{itemize}

\end{frame}

\begin{frame}{The Algorithm}

    \begin{itemize}
        \item A variable for every pattern vertex.
        \item Domains range over target vertices.
        \item Clever filtering at the top of search using neighbourhood degree sequences and paths.
        \item Bit-parallel forward checking search:
            \begin{itemize}
                \item Adjacency and pre-computed path count constraints.
                \item Cheap hacky all-different propagator.
                \item Smallest domain first, tie-breaking on highest degree.
            \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}{Is It Any Good?}

    \includegraphics<1>{gen-graph-others.pdf}%
    \includegraphics<2>{gen-graph-others-zoom.pdf}

\end{frame}

\begin{frame}{Value-Ordering Heuristics}
    \begin{itemize}
        \item Highest static degree first.
        \item But degree spread is usually quite low\ldots
    \end{itemize}
\end{frame}

\begin{frame}{Sanity Check}

    \includegraphics<1>{gen-graph-value-ordering-heuristics.pdf}%
    \includegraphics<2>{gen-graph-value-ordering-heuristics-unsat.pdf}%

\end{frame}

\begin{frame}{Discrepancy Search}

    \includegraphics<1>{gen-graph-scatter-dds.pdf}

\end{frame}

\begin{frame}{Restarts and Nogoods}

    \begin{itemize}
        \item Restarts to affect \emph{value}-ordering heuristics.
        \item Two watched literals on decision nogoods to avoid repeating ourselves.
        \item Luby with a magic constant multiplier.
            \begin{itemize}
                \item 666, chosen by divine revelation.
            \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}{Random Value-Ordering with Restarts}

    \includegraphics<1>{gen-graph-random-restarts.pdf}%
    \includegraphics<2>{gen-graph-scatter-random.pdf}

\end{frame}

\begin{frame}{Biased Value-Ordering}

    \only<1>{
    \begin{itemize}
        \item Select a vertex $v'$ from the chosen domain $D_v$ with
            probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]
        \item Looks a lot like $\mathit{softmax}$, which uses base $e$.
    \end{itemize}
    }%
    \includegraphics<2>{gen-graph-value-ordering-heuristics-biased.pdf}%
    \includegraphics<3>{gen-graph-scatter-biased.pdf}

\end{frame}

\begin{frame}{Biased Value-Ordering with Restarts}

    \includegraphics<1>{gen-graph-restarts.pdf}%
    \includegraphics<2>{gen-graph-scatter-final.pdf}

\end{frame}

\section{Maximum Common Subgraph}

\begin{frame}{Maximum Common Subgraph Algorithms}

    \includegraphics<1>{gen-graph-mcs.pdf}%
    \includegraphics<2>{gen-graph-scatter-kdown.pdf}%
    \includegraphics<3>{gen-graph-scatter-mcsplit.pdf}%

\end{frame}

\section{Parallel}

\begin{frame}{Parallel}

    \includegraphics<1>{gen-graph-parallel.pdf}%
    \includegraphics<2>{gen-graph-as-parallel.pdf}%
    \includegraphics<3>{gen-graph-scatter-parallel.pdf}%

\end{frame}

\begin{frame}{Properties of This Sort of Parallel Search}
    \begin{itemize}
        \item Reproducible (assuming each thread has its own deterministic random number generator)
        \item Can perform worse than sequential on satisfiable instances:
            \begin{itemize}
                \item Nogoods can knock out values from domains, leading to a different random
                    search order, even if the first thread uses the same seed.
                \item Challenge for you: design a sane and fast biased random selection algorithm
                    that is immune to this, and that is statistically the same sequentially. It
                    feels like this should be possible\ldots
                \item Also, our all different propagator isn't monotonic or idempotent.
            \end{itemize}

        \item Similarly, adding extra threads can make it worse.
    \end{itemize}
\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{center}
        \vspace*{4em}
        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
        \vspace{1em}
        \href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}} \\
    \end{center}

    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };
    \end{tikzpicture}
\end{frame}

>>>>

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}
\usepackage{listings}
\usepackage{csquotes}

\RequirePackage[tt=false, type1=true]{libertine}
\RequirePackage[varqu]{zi4}
\RequirePackage[libertine]{newtxmath}
\RequirePackage[T1]{fontenc}

@@

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.4cm)}] at (current page.north east) {
            \begin{minipage}{0.20\paperwidth}
                \begin{center}
                \includegraphics[keepaspectratio=true,scale=0.35]{UoG_keyline.pdf} \\[0.15cm]
                \includegraphics[keepaspectratio=true,scale=0.8]{UStA_Logo_SP.eps}
                \end{center}
            \end{minipage}
        };

        \node [anchor=west, xshift=0.2cm, yshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

@@

\makeatletter
\newenvironment{justborderframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \begin{frame}[#1]{}
    }{
    \end{frame}
}
\makeatother

@@

\author{Blair Archibald \and Fraser Dunlop \and Ruth Hoffmann \and Ciaran McCreesh \and Patrick
Prosser \and James Trimble}
\title{Sequential and Parallel Solution-Biased Search for Subgraph Algorithms}

@@

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, height=\paperheight]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Subgraph Isomorphism}
    \centering
    \begin{tikzpicture}
        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {\vphantom{1}};
        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {\vphantom{1}};
        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {\vphantom{1}};
        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {\vphantom{1}};

        \draw [ultra thick] (Na) -- (Nb);
        \draw [ultra thick] (Nb) -- (Nc);
        \draw [ultra thick] (Nc) -- (Nd);
        \draw [ultra thick] (Nb) -- (Nd);

        \node [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {\vphantom{1}};
        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {\vphantom{1}};
        \node [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {\vphantom{1}};
        \node [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {\vphantom{1}};
        \node [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {\vphantom{1}};
        \node [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {\vphantom{1}};

        \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw [ultra thick] (N1) -- (N3);
        \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw [ultra thick] (N3) -- (N5);
        \draw [ultra thick] (N3) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw [ultra thick] (N5) -- (N6);
        \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}
\end{frame}

\begin{frame}{Subgraph Finding, as a Constraint Program}
    \begin{itemize}
        \item A variable for each pattern vertex. The domains are all of the target vertices.
        \item At least two sets of constraints:
            \begin{itemize}
                \item Adjacent pairs of vertices must be mapped to adjacent pairs of vertices.
                \item All different.
            \end{itemize}
        \item Then we get clever:
            \begin{itemize}
                \item Extra constraints about degrees, paths, \ldots
                \item Very good variable- and value-ordering heuristics
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{The Glasgow Subgraph Solver}
    \begin{center}
        \url{https://github.com/ciaranm/glasgow-subgraph-solver}
    \end{center}

    \begin{itemize}
        \item Subgraph isomorphism, and all its variants (induced / non-induced, homomorphism,
            locally injective, labels, side constraints, directed, \ldots).
        \item Also special algorithms for clique.
    \end{itemize}
\end{frame}

\begin{frame}{Value-Ordering Heuristics}
    \begin{itemize}
        \item Largest target degree first.
    \end{itemize}
\end{frame}

\begin{frame}{Sanity Check}
    \begin{center}
        \includegraphics<1>{gen-graph-value-ordering.pdf}%
        \includegraphics<2>{gen-graph-value-ordering-unsat.pdf}
    \end{center}
\end{frame}

\begin{frame}{However\ldots}
    \begin{itemize}
        \item What if several vertices have the same degree?
        \item Is a vertex of degree $10$ really that much better than a vertex of degree $9$?
    \end{itemize}
\end{frame}

\begin{frame}{Discrepancy Search?}
    \begin{center}
        \includegraphics<1>{gen-graph-value-ordering-dds.pdf}%
        \includegraphics<2>{gen-graph-value-ordering-dds-unsat.pdf}%
        \includegraphics<3>{gen-graph-value-ordering-dds-scatter.pdf}
    \end{center}
\end{frame}

\begin{frame}{Value-Ordering Heuristics as Distributions}
    \begin{itemize}
        \item Traditional view: value-ordering defines a search order.
        \item New view: value-ordering defines \textcolor{uofgcobalt}{what proportion of the search
            effort} should be spent on different subproblems.
    \end{itemize}
\end{frame}

\begin{frame}{A Slightly Random Value-Ordering Heuristic}
    \only<1>{
    \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \]
    \begin{itemize}
        \item Equally likely to pick between two vertices of degree $d$.
        \item Twice as likely to select a vertex of degree $d$ than a vertex of degree $d - 1$.
        \item In the paper: \textcolor{uofgcobalt}{solution density} and expected distribution of solutions.
    \end{itemize}}\only<2>{
    \begin{center}
        \includegraphics<2>{gen-graph-bias-scatter.pdf}%
    \end{center}}
\end{frame}

\begin{frame}{Restarts and Nogood Recording}
    \begin{itemize}
        \item Aggressive restarts: every 100ms.
        \item Nogood recording and 2WL to avoid repeating work.
    \end{itemize}
\end{frame}

\begin{frame}{Is It Better?}
    \begin{center}
        \includegraphics<1>{gen-graph-sbs.pdf}%
        \includegraphics<2>{gen-graph-sbs-unsat.pdf}%
        \includegraphics<3>{gen-graph-sbs-scatter.pdf}%
    \end{center}
\end{frame}

\begin{frame}{Parallel Search}
    \begin{itemize}
        \item Each thread gets its own random seed.
        \item Barrier synchronise on restarts.
        \item Share nogoods.
    \end{itemize}
\end{frame}

\begin{frame}{Is It Even Betterer?}
    \includegraphics<1>{gen-graph-par-scatter.pdf}%
    \includegraphics<2>{gen-graph-par.pdf}%
    \includegraphics<3>{gen-graph-dist.pdf}%
\end{frame}

\begin{frame}{Broader Perspective}
    \begin{itemize}
        \item Also good for maximum common subgraph algorithms.
        \item Does this work for CP in general?
        \item Can we finally kill off vanilla sequential backtracking search?
    \end{itemize}
\end{frame}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner
        sep=0pt]{\includegraphics[keepaspectratio=true, width=\paperwidth]{background2.jpg}};
    }

    \begin{frame}[plain,noframenumbering]
        \begin{tikzpicture}[remember picture, overlay]
            \node at (current page.north west) {
                \begin{tikzpicture}[remember picture, overlay]
                    \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
                \end{tikzpicture}
            };

            \node [anchor=north west, shift={(0.6cm,-0.8cm)}] at (current page.north west) {
                \textcolor{white}{\url{http://www.dcs.gla.ac.uk/~ciaran}}
            };

            \node [anchor=north west, shift={(0.6cm,-1.4cm)}] at (current page.north west) {
                    \textcolor{white}{\href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}}}
            };
        \end{tikzpicture}
    \end{frame}
}
